This page and the next page should be removed

### 2021-04-05 analysis/manuscript.Rmd
add stan and tfr try-outs

In the two experiments subsequently reported, participants were presented with real-word objects that were presumed to be unfamiliar to most of them. They first viewed each of these objects without any semantic information. This first part served as a naive baseline to rule out that ERPs would differ in response to the objects based solely on low-level visual differences. Next, participants viewed each unfamiliar object for a second time, now preceded by verbal keywords. These keywords could either be matching the typical function of the object, thus making it possible for participants to understand what kind of object they were viewing, or they could be non-matching by describing the function of a different object, thus keeping the perception of the object semantically naive. During this second part, we were able to measure the online influence of semantic insight on object-evoked ERPs as it happened. Finally, the objects were presented for a third time (without keywords, as in the first part) to investigate downstream effects of having acquired semantic knowledge about them---thereby mimicking previous studies [e.g., @abdelrahman2008; @samaha2018; @weller2019]. In all three parts, we examined the influence of semantic information on ERPs associated with lower-level visual perception (P1 component, 100--150 ms), higher-level visual perception (N170 component, 150--200 ms), and semantic processing (N400 component, 400--700 ms).

@@

>>>>

In the two experiments subsequently reported, participants were presented with real-world objects that were presumed to be unfamiliar to most of them. They first viewed each of these objects without any semantic information. This first part served as a naive baseline to rule out that ERPs would differ in response to the objects based solely on low-level visual differences. Next, participants viewed each unfamiliar object for a second time, now preceded by verbal keywords. These keywords could either be matching the typical function of the object, thus making it possible for participants to understand what kind of object they were viewing, or they could be non-matching by describing the function of a different object, thus keeping the perception of the object semantically naive. During this second part, we were able to measure the online influence of semantic insight on object-evoked ERPs as it happened. Finally, the objects were presented for a third time (without keywords, as in the first part) to investigate downstream effects of having acquired semantic knowledge about them---thereby mimicking previous studies [e.g., @abdelrahman2008; @samaha2018; @weller2019]. In all three parts, we examined the influence of semantic information on ERPs associated with lower-level visual perception (P1 component, 100--150 ms), higher-level visual perception (N170 component, 150--200 ms), and semantic processing (N400 component, 400--700 ms).

@@

```{r bayes, eval=FALSE}
library(brms)
options(mc.cores = 6)

data_joint_subs %>%
  filter(condition %in% c("Informed", "Naive")) %>%
  droplevels() %>%
  select(part, condition, participant, item, P1, N170, N400) %>%
  na.omit() %>%
  as.list() -> data_brm

contrasts_condition <- t(cbind(c("Informed" = 1, "Naive" = -1)))
contrasts_part <- t(cbind(c("I" = -1, "II" = 1, "III" = 0), c("I" = 0, "II" = -1, "III" = 1)))
contrasts(data_brm$condition) <- MASS::ginv(contrasts_condition)
contrasts(data_brm$part) <- MASS::ginv(contrasts_part)

formula <- P1 ~ part / condition + (part / condition | participant) + (part / condition | item)
prior <- set_prior("cauchy(0, 0.707)", class = "b")

mod_brm <- brm(
  formula = formula,
  data = data_brm,
  family = gaussian(),
  prior = prior,
  sample_prior = "yes",
  save_pars = save_pars(all = TRUE),
  chains = 4,
  iter = 20000,
  warmup = 2000,
  cores = 4,
  control = list(adapt_delta = 0.9)
)

# mod_prior <- brm(formula = formula, data = data_brm, prior = prior, sample_prior = "only")
# prior_summary(mod_prior)
# mcmc_plot(mod_prior)
#
# mod_post <- update(mod_prior, sample_prior = "yes")
# summary(mod_post)
# mcmc_plot(mod_post, pars = "b")
# plot(mod_post)

hypothesis <- paste0("part", c("I", "II", "III"), ":condition1 = 0")
bf_brms <- hypothesis(mod_post, hypothesis = hypothesis)
bf_brms$hypothesis$bf10 <- 1 / bf_brms$hypothesis$Evid.Ratio
bf_brms

bf_bayestest <- bayestestR::bayesfactor(mod_post)
bf_bayestest

mod_brm_s <- brm(
  bf(P1 ~ part / condition + (part / condition | participant) + (part / condition | item),
     sigma ~ 1 + (1 | participant)),
  prior = c(
    prior(cauchy(0, 0.707), class = b),
    prior(student_t(3, 0, 5.7), class = Intercept, dpar = sigma),
    prior(normal(0, 5), class = sd, group = participant, dpar = sigma)),
  data = data_brm,
  family = gaussian(),
  sample_prior = "yes",
  save_pars = save_pars(all = TRUE),
  chains = 4,
  iter = 20000,
  warmup = 2000,
  cores = 4,
  control = list(adapt_delta = 0.9)
)
```

```{python tfr, eval=FALSE}
import mne
import numpy as np
#import matplotlib.pyplot as plt

# epochs = preproc(
#   vhdr_fname="data/exp1/eeg/Vp0004.vhdr",
#   metadata=r.data_exp1.query("participant == 'VP_04'"),
#   n_components=15,
#   method="fastica",
#   l_freq=None,
#   h_freq=None,
#   event_id=dict(match=221, mismatch=222),
#   tmin=-0.5,
#   tmax=1.498,
#   baseline=(-0.2, 0),
#   reject=dict(eeg=200)
# )
# epochs.save("analysis/export/vp04-epo.fif")

epochs = mne.read_epochs("analysis/export/unf-exp1-epo.fif")

evokeds = compute_evokeds(epochs_c, grand=True, data_only=False)
evokeds_i = evokeds["II"]["Informed"]
evokeds_n = evokeds["II"]["Naive"]

freqs = np.logspace(*np.log10([6, 35]), num=8)
n_cycles = freqs / 2.
power_i = mne.time_frequency.tfr_morlet(evokeds_i, freqs=freqs, n_cycles=n_cycles, use_fft=True, return_itc=False, average=True, n_jobs=1)
power_n = mne.time_frequency.tfr_morlet(evokeds_n, freqs=freqs, n_cycles=n_cycles, use_fft=True, return_itc=False, average=True, n_jobs=1)
power_i = power_i.apply_baseline(mode='mean', baseline=(-.400, -.100))
power_n = power_n.apply_baseline(mode='mean', baseline=(-.400, -.100))

power_d = power_i.copy()
power_d.data = power_i.data - power_n.data

power_d.plot()

# epochs.resample(250., npad='auto')

roi = r.comps["roi"][0] + r.comps["roi"][1]
idxs = np.where(np.in1d(epochs.ch_names, roi))

epochs_i = epochs["part == 'II' & condition == 'Informed'"].pick_channels(roi)
epochs_n = epochs["part == 'II' & condition == 'Naive'"].pick_channels(roi)

epochs_c = mne.channels.combine_channels(epochs, groups=dict(avg=range(62)))

epochs_c.metadata = epochs.metadata
epochs_s = epochs_c["part == 'II'"]
# epochs_i = epochs_s["condition == 'Informed'"]
# epochs_n = epochs_s["condition == 'Naive'"]

# power_i = mne.time_frequency.tfr_morlet(epochs_i, freqs=freqs, n_cycles=n_cycles, use_fft=True, return_itc=False, average=True, n_jobs=1)
# power_n = mne.time_frequency.tfr_morlet(epochs_n, freqs=freqs, n_cycles=n_cycles, use_fft=True, return_itc=False, average=True, n_jobs=1)
# power_d = power_i - power_n
# 
# power_d.plot([0], baseline=(-0.5, 0), mode='percent', vmin=-1, vmax=1)

# define frequencies of interest (log-spaced)
#freqs = np.logspace(*np.log10([6, 35]), num=8)
freqs = np.arange(3, 50, 1.6)
#n_cycles = freqs / 2.  # different number of cycle per frequency
n_cycles = np.linspace(3, 10, len(freqs))
power = mne.time_frequency.tfr_morlet(epochs_c, freqs=freqs, n_cycles=n_cycles, use_fft=True, return_itc=False, average=True, n_jobs=1)

power.plot([0], baseline=(-0.5, 0), mode='percent')

power.plot_topo(baseline=(-0.5, 0), mode='logratio')
power.plot_topomap(ch_type='eeg', tmin=0.15, tmax=0.2, fmin=8, fmax=12, baseline=(-0.5, 0), mode='logratio', title='Alpha')
power.plot_topomap(ch_type='eeg', tmin=0.15, tmax=0.2, fmin=13, fmax=25, baseline=(-0.5, 0), mode='logratio', title='Beta')

import numpy as np
import matplotlib.pyplot as plt
from mne.baseline import rescale
from mne.stats import bootstrap_confidence_interval
iter_freqs = [
    ('Theta', 4, 7),
    ('Alpha', 8, 12),
    ('Beta', 13, 25),
    ('Gamma', 30, 45)
]
raw_fname = "data/exp1/EEG/Vp0004.vhdr"
tmin, tmax = -0.5, 1.498
baseline = None
frequency_map = list()
event_id = dict(match=222)
for band, fmin, fmax in iter_freqs:
    # (re)load the data to save memory
    raw = mne.io.read_raw_brainvision(raw_fname, scale=1e6, preload=True)
    raw.filter(fmin, fmax, n_jobs=1,  # use more jobs to speed up.
               l_trans_bandwidth=1,  # make sure filter params are the same
               h_trans_bandwidth=1)  # in each band and skip "auto" option.
    events, _ = mne.events_from_annotations(raw, verbose=False)
    epochs = mne.Epochs(raw, events, event_id, tmin, tmax, baseline=baseline,
                        reject=dict(eeg=200), preload=True)
    # remove evoked response
    epochs.subtract_evoked()
    # get analytic signal (envelope)
    epochs.apply_hilbert(envelope=True)
    frequency_map.append(((band, fmin, fmax), epochs.average()))
    del epochs
del raw
# Helper function for plotting spread
def stat_fun(x):
    """Return sum of squares."""
    return np.sum(x ** 2, axis=0)
# Plot
fig, axes = plt.subplots(4, 1, figsize=(10, 7), sharex=True, sharey=True)
colors = plt.get_cmap('winter_r')(np.linspace(0, 1, 4))
for ((freq_name, fmin, fmax), average), color, ax in zip(
        frequency_map, colors, axes.ravel()[::-1]):
    times = average.times * 1e3
    gfp = np.sum(average.data ** 2, axis=0)
    gfp = mne.baseline.rescale(gfp, times, baseline=(None, 0))
    ax.plot(times, gfp, label=freq_name, color=color, linewidth=2.5)
    ax.axhline(0, linestyle='--', color='grey', linewidth=2)
    ci_low, ci_up = bootstrap_confidence_interval(average.data, random_state=0,
                                                  stat_fun=stat_fun)
    ci_low = rescale(ci_low, average.times, baseline=(None, 0))
    ci_up = rescale(ci_up, average.times, baseline=(None, 0))
    ax.fill_between(times, gfp + ci_up, gfp - ci_low, color=color, alpha=0.3)
    ax.grid(True)
    ax.set_ylabel('GFP')
    ax.annotate('%s (%d-%dHz)' % (freq_name, fmin, fmax),
                xy=(0.95, 0.8),
                horizontalalignment='right',
                xycoords='axes fraction')
axes.ravel()[-1].set_xlabel('Time [ms]')
plt.show()
```
